---
title: "US Economic Outlook"
author: "satrio"
date: "2025-09-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(patchwork)
library(gt)
library(gtExtras)
library(ollamar)
library(jsonlite)

# parameters
cutoff_date <- as.character(ymd(Sys.Date()))

# fetch data
source("~/Economic_outlook/data_fetch.R")

# graph and table script
source("~/Economic_outlook/real_economy/data_viz_real_economy.R")

# Download Real economy time series data
real_economy_id <- c("GDP", "GDPC1",
                     "INDPRO", "IPMAN", "IPDMAN", "IPNMAN", "IPMINE", "IPUTIL",
                     "RSXFS", "RSAFS", "RSMVPD")

real_economy_data <- pull_data_fred(series_vector = real_economy_id, start_date = "2000-01-30", end_date = "2025-09-22")

# gdp components
gdp_components_id <- c("PCECC96", "DGDSRX1Q020SBEA", "PCESVC96",
                       "GPDIC1", "CBIC1", "FPIC1",
                       "NETEXC", "EXPGSC1", "IMPGSC1",
                       "GCEC1", 
                       "A960RX1Q020SBEA")

gdp_components <- pull_data_fred(series_vector = gdp_components_id, start_date = "2001-01-30", end_date = "2025-09-22")

# labour market 
labour_market_id <- c("UNRATE", "PAYEMS", "LNS11300060", "JTSJOL", "AWHMAN", "CES0500000003")

labour_market_data <- pull_data_fred(series_vector = labour_market_id, start_date = "2000-01-30", end_date = "2025-09-22")


options(scipen = 999)
```

## Executive Summary:

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

## Real Economy Outlook

### GDP Analysis

```{r first paragraph}
# Real GDP analysis
prompt_gdp_analysis <- "You are a expert macroeconomist and you are tasked to
                        generate a report on US economy. The first thing you will need
                        to do is analysing US GDP and the GDP section is divided into
                        5 paragraph in total. Right now, you are in the first paragraph and
                        you required to explain why GDP analysis is important for this economic report.
                        generate paragraph consisting 100 maximum words. Please return only the paragpraph, 
                        no title and omit the first sentence like 'Okay, here's the first'."

resp_gdp_analysis <- generate("gemma3:1b", prompt_gdp_analysis, output = "text")

```

`r resp_gdp_analysis`

```{r second paragraph}
gdp_json_second_par <- gdp_line_chart_data_ai(data = real_economy_data,
                                              cutoff = "2018-01-01",
                                              type = "yoy_growth")

prompt_gdp_analysis_second <- paste0("You are an expert macroeconomist and you are tasked to generate a report on US economy. The first thing you will need to do is analysing US GDP and the GDP section is divided into 5 paragraph in total. Right now, you are in the second paragraph and you are tasked to generate paragraph consisting 200 maximum words about the Nominal and Real GDP Growth rate movements across time, also provide a quick analysis on movements and direction of the GDP growth rate. Please return only the paragpraph, no title, omit the first sentence like 'Okay, here's the first' and please analyse the given JSON data only and omit paragraph title, future projection and key take aways. Here is the data in JSON format:", gdp_json_second_par)

resp_gd_second_par <- generate("gemma3:1b", prompt_gdp_analysis_second, output = "text")

```

#### Real & Nominal GDP level and Growth Rate

```{r Real GDP QoQ and YoY growth rate, echo=FALSE, fig.width=15, fig.height=8}
yoy_gdp_line <- gdp_line_chart(data = real_economy_data,
                               cutoff = "2018-01-01",
                               type = "yoy_growth")

qoq_gdp_line <- gdp_line_chart(data = real_economy_data,
                               cutoff = "2018-01-01",
                               type = "qoq_annualised")

yoy_gdp_line
```

`r resp_gd_second_par`

#### Nominal vs Real GDP table

```{r Nominal vs Real GDP table, echo=FALSE}
gdp_table_func(data = real_economy_data, cutoff = "2018-01-01", type = "growth")
```

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

#### Real GDP Component Driver

```{r Real GDP Component, echo=FALSE}
chart_gdp_comp <- gdp_component_chart(data1 = gdp_components, data2 = real_economy_data, cutoff = "2018-01-01")
chart_gdp_comp$contrib_comp
```

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

#### Real GDP Contribution Table

```{r Real GDP Contribution Table, echo=FALSE}
gdp_contrib_table(data1 = gdp_components, data2 = real_economy_data, cutoff = "2018-01-01")
```

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

```{r testing markdown}
resp <- generate("gemma3:1b", "Could give a summarise on mining outlook in Indonesia?", output = "text")

```

`r resp`
