---
title: "US Economic Outlook"
author: "satrio"
date: "2025-09-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(patchwork)
library(gt)
library(gtExtras)

# parameters
cutoff_date <- as.character(ymd(Sys.Date()))

# fetch data
source("~/Economic_outlook/data_fetch.R")

# graph and table script
source("~/Economic_outlook/real_economy/data_viz_real_economy.R")

# Download Real economy time series data
real_economy_id <- c("GDP", "GDPC1",
                     "INDPRO", "IPMAN", "IPDMAN", "IPNMAN", "IPMINE", "IPUTIL",
                     "RSXFS", "RSAFS", "RSMVPD")

real_economy_data <- pull_data_fred(series_vector = real_economy_id, start_date = "2000-01-30", end_date = "2025-09-22")

# gdp components
gdp_components_id <- c("PCECC96", "DGDSRX1Q020SBEA", "PCESVC96",
                       "GPDIC1", "CBIC1", "FPIC1",
                       "NETEXC", "EXPGSC1", "IMPGSC1",
                       "GCEC1", 
                       "A960RX1Q020SBEA")

gdp_components <- pull_data_fred(series_vector = gdp_components_id, start_date = "2001-01-30", end_date = "2025-09-22")

# labour market 
labour_market_id <- c("UNRATE", "PAYEMS", "LNS11300060", "JTSJOL", "AWHMAN", "CES0500000003")

labour_market_data <- pull_data_fred(series_vector = labour_market_id, start_date = "2000-01-30", end_date = "2025-09-22")


options(scipen = 999)
```

## Executive Summary:

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

-   Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

## Real Economy Outlook

### GDP Analysis

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

#### Real & Nominal GDP level and Growth Rate

```{r Real GDP QoQ and YoY growth rate, echo=FALSE, fig.width=15, fig.height=8}
yoy_gdp_line <- gdp_line_chart(data = real_economy_data,
                               cutoff = "2018-01-01",
                               type = "yoy_growth")

qoq_gdp_line <- gdp_line_chart(data = real_economy_data,
                               cutoff = "2018-01-01",
                               type = "qoq_annualised")

qoq_gdp_line + yoy_gdp_line
```

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

#### Nominal vs Real GDP table

```{r Nominal vs Real GDP table, echo=FALSE}
gdp_table_func(data = real_economy_data, cutoff = "2018-01-01", type = "growth")
```

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

#### Real GDP Component Driver

```{r Real GDP Component, echo=FALSE}
chart_gdp_comp <- gdp_component_chart(data1 = gdp_components, data2 = real_economy_data, cutoff = "2018-01-01")
chart_gdp_comp$contrib_comp
```

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

#### Real GDP Contribution Table

```{r Real GDP Contribution Table, echo=FALSE}
gdp_contrib_table(data1 = gdp_components, data2 = real_economy_data, cutoff = "2018-01-01")
```

It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.
